### 2025-11-23 article/sections/abstract.tex
article tex seperated. Graph construction part added

\section{Abstract}

tbd

% $$\hat{r} = \frac{N}{E}$$

% $$LL_{\lambda} = \frac{1}{2}\chi^2_{0.025, 2N}$$
% $$UL_{\lambda} = \frac{1}{2}\chi^2_{0.975, 2N + 2}$$

% $$LL_{r} = LL_{\lambda} / E$$
% $$UL_{r} = UL_{\lambda} / E$$