### 2025-11-23 article/sections/defining_traffic_scene_graph.tex

\section{Defining a traffic scene graph}
\label{chapter:defining_a_traffic_scene_graph}

still to be added: \url{https://sagroups.ieee.org/adwg/wp-content/uploads/sites/661/2024/10/ADWG_STV2_whitepaper.pdf}

still to be added: \url{https://www.vvm-projekt.de/securedl/sdl-eyJ0eXAiOiJKV1QiLCJhbGciOiJIUzI1NiJ9.eyJpYXQiOjE3NDc2OTIxNjgsImV4cCI6MTc0Nzc4MjE2OCwidXNlciI6MCwiZ3JvdXBzIjpbMCwtMV0sImZpbGUiOiJmaWxlYWRtaW4vdXNlcl91cGxvYWQvVlZNZXRob2Rlbl9adXNhbW1lbmZhc3NlbmRlcl9BYnNjaGx1c3NiZXJpY2h0X3YxLjBfenVyX1Zlc_f9d402f084002b57/VVMethoden_Zusammenfassender_Abschlussbericht_v1.0_zur_Veroeffentlichung.pdf}

\cite{newman2010networks}

\cite{bagschik2018ontologybasedscenecreation}

\cite{fremont2020scenic}

\subsection{Map Graph Construction}
\label{subsec:map_graph}

The map graph serves as the foundational structure for traffic scene graph construction, encoding the 
spatial relationships between lanes in the road network. It is a directed multigraph 
$G_{\text{map}} = (V_{\text{map}}, E_{\text{map}})$ where nodes represent individual lanes and edges 
encode three fundamental spatial relationships between lanes.

Each node $v \in V_{\text{map}}$ represents a lane segment and stores geometric and semantic 
information including:
\begin{itemize}
    \item Lane boundaries (left and right boundaries as polylines)
    \item Lane length
    \item Lane centerline
    \item Intersection status (whether the lane is within an intersection)
    \item Road type and lane type information
\end{itemize}

The map graph defines three edge types that capture different spatial relationships:

\begin{enumerate}
    \item \textbf{Following edges} ($e_{\text{following}}$): Connect lanes that form a continuous 
          path in the same direction.

    \item \textbf{Neighbor edges} ($e_{\text{neighbor}}$): Connect adjacent lanes traveling in 
          the same direction.

    \item \textbf{Opposite edges} ($e_{\text{opposite}}$): Connect lanes traveling in opposite 
          directions. 
\end{enumerate}

The map graph construction algorithm processes the original map data from Argoverse or CARLA  
to extract these relationships. For intersection detection, lanes are analyzed for geometric 
overlap, and all overlapping lanes are marked as intersection lanes to ensure proper handling of 
complex road geometries.

%TODO: add figure

\subsection{Actor Graph Construction}
\label{subsec:actor_graph}

The actor graph $G_{\text{actor}} = (V_{\text{actor}}, E_{\text{actor}})$ represents the dynamic 
relationships between vehicles and other actors in a traffic scene at a specific timestep. This 
graph is constructed using a two-phase algorithm that separates relation discovery from graph 
construction. The second step serves to reduce the number of relations in the graph for computational tractability, 
and enables the representation of relationships between actors through indirect paths involving multiple intermediate nodes.

Each node $v \in V_{\text{actor}}$ represents an actor (vehicle, pedestrian, etc.) and stores the 
following attributes:

\begin{itemize}
    \item \textbf{Primary lane ID}: The lane on which the actor is primarily located
    \item \textbf{Lane IDs}: List of all lanes the actor occupies (for actors spanning multiple 
          lanes)
    \item \textbf{Longitudinal position} ($s$): Position along the lane's centerline
    \item \textbf{3D position} ($x, y, z$): Cartesian coordinates
    \item \textbf{Longitudinal speed}: Speed along the lane direction
    \item \textbf{Actor type}: Classification (vehicle, pedestrian, etc.)
    \item \textbf{Lane change indicator}: Boolean flag indicating if the actor changed lanes from 
          the previous timestep
\end{itemize}

Note that the node structure is extensible---users can add additional attributes such as 
acceleration, heading angle, or vehicle dimensions as needed for their coverage model.

Each edge $e \in E_{\text{actor}}$ represents a relationship between two actors and contains:

\begin{itemize}
    \item \textbf{Edge type}: One of four relation types between actors (leading, following, neighbor, opposite)
    \item \textbf{Path length}: Distance in meters along the lane network 
\end{itemize}

Similar to nodes, edge attributes can be extended with additional information such as relative 
velocities, time-to-collision, or interaction probabilities.

\label{subsubsec:relation_types}

The actor graph defines four types of relationships between actors, ordered by semantic hierarchy:

\begin{enumerate}
    \item \textbf{Following/Leading} (\texttt{following\_lead}, \texttt{leading\_vehicle}): 
          Represents longitudinal relationships where one actor follows another in the same 
          direction and lane.

    \item \textbf{Neighbor} (\texttt{neighbor\_vehicle}): Represents lateral relationships 
          between actors on lanes traveling in the same direction. The actors need not be on 
          immediately adjacent lanes.

    \item \textbf{Opposite} (\texttt{opposite\_vehicle}): Represents relationships between actors 
          on lanes traveling in opposite directions. Similar to neighbor relationships, the actors 
          need not be on immediately opposite lanes.
\end{enumerate}

%TODO: add figure

\subsection{Hierarchical Graph Construction Algorithm}
\label{subsec:construction_algorithm}

The actor graph construction follows a two-phase approach that separates relation discovery from 
graph building, enabling hierarchical processing and preventing redundant edge creation.

\subsubsection{Phase 1: Relation Discovery}
\label{subsubsec:discovery_phase}

The discovery phase identifies all potential relationships between actor pairs based on distance 
constraints and map graph structure. For each pair of actors $(A, B)$ at timestep $t$, the 
algorithm determines their primary lanes and checks if a path exists in the map graph connecting 
them. The path structure determines the relationship type:

\begin{itemize}
    \item \textbf{Following/Leading}: If both actors are on the same lane, or if the path consists 
          entirely of following edges.

    \item \textbf{Neighbor}: If the path contains exactly one neighbor edge and the remaining edges 
          are following edges. Note that this does not require the actors to be on immediately 
          adjacent lanes---the path may contain multiple following edges before and after the single 
          neighbor edge.

    \item \textbf{Opposite}: If the path contains exactly one opposite edge and the remaining edges 
          are following edges. Similar to neighbor relationships, this does not require the actors 
          to be on immediately opposite lanes.
\end{itemize}

Each relationship is validated using two distance checks: the lane-based path length ensures that 
curved roads are properly accounted for, while the Euclidean distance ensures that only actors in 
close proximity are considered. All discovered relationships are stored in a relations dictionary 
for the construction phase.

%TODO: add figure

\subsubsection{Phase 2: Hierarchical Graph Construction}
\label{subsubsec:construction_phase}

The construction phase builds the actor graph by adding edges in a hierarchical order, ensuring 
that higher-priority relationships are established first and redundant edges are prevented.

\paragraph{Construction Order}

Edges are added in three stages, following the semantic hierarchy:

\begin{enumerate}
    \item \textbf{Leading/Following edges} (highest priority): All following/leading relationships 
          are processed first, sorted by path length (shortest first).

    \item \textbf{Neighbor edges} (medium priority): Neighbor relationships are processed second, 
          sorted by absolute path length.

    \item \textbf{Opposite edges} (lowest priority): Opposite relationships are processed last, 
          sorted by absolute path length.
\end{enumerate}

\paragraph{Redundancy Prevention}

Before adding an edge between actors $A$ and $B$, the algorithm checks if a path already exists 
in the current actor graph with length $\leq$ the maximum node distance for that relation type. 
This prevents redundant edges that would create triangles or longer indirect paths when shorter 
direct paths exist.

For leading/following and neighbor relationships, each direction ($A \to B$ and $B \to A$) is 
checked independently, as these relationships can be asymmetric. For opposite relationships, if 
either direction has an existing path, both directions are rejected, as opposite relationships are 
inherently symmetric.

The path checking uses breadth-first search to find the shortest path in the actor graph, counting 
the number of edges (not nodes) in the path. This ensures that if a path exists with $k$ edges 
where $k \leq \text{max\_node\_distance}$, the direct edge is not added.

The graph is updated immediately after each edge addition, ensuring that path checks for 
subsequent edges use the current graph state. This incremental approach guarantees that the 
hierarchical construction works correctly and prevents race conditions.

%TODO: add figure

\subsubsection{Input Parameters}
\label{subsubsec:parameters}

Table~\ref{tab:construction_parameters} summarizes all input parameters for the actor graph 
construction algorithm, their descriptions, and the values used in our experiments.

\begin{table}[h]
\centering
\caption{Input parameters for actor graph construction}
\label{tab:construction_parameters}
\begin{tabular}{llp{6cm}l}
\hline
\textbf{Parameter} & \textbf{Type} & \textbf{Description} & \textbf{Value} \\
\hline
\multicolumn{4}{l}{\textit{Distance limits (discovery phase)}} \\
\hline
\texttt{max\_distance\_lead\_veh\_m} & float & Maximum distance in meters for leading/following 
relationships & 100 \\
\texttt{max\_distance\_neighbor\_forward\_m} & float & Maximum distance in meters for forward 
neighbor relationships & 50 \\
\texttt{max\_distance\_neighbor\_backward\_m} & float & Maximum distance in meters for backward 
neighbor relationships & 50 \\
\texttt{max\_distance\_opposite\_forward\_m} & float & Maximum distance in meters for forward 
opposite relationships & 100 \\
\texttt{max\_distance\_opposite\_backward\_m} & float & Maximum distance in meters for backward 
opposite relationships & 10 \\
\hline
\multicolumn{4}{l}{\textit{Node distance limits (construction phase)}} \\
\hline
\texttt{max\_node\_distance\_leading} & int & Maximum number of edges in actor graph path for 
leading/following & 3 \\
\texttt{max\_node\_distance\_neighbor} & int & Maximum number of edges in actor graph path for 
neighbor & 2 \\
\texttt{max\_node\_distance\_opposite} & int & Maximum number of edges in actor graph path for 
opposite & 2 \\
\hline
\multicolumn{4}{l}{\textit{Timestep configuration}} \\
\hline
\texttt{delta\_timestep\_s} & float & Time step increment in seconds for temporal graph 
creation & 1.0 \\
\hline
\end{tabular}
\end{table}

The distance limits in the discovery phase determine which potential relationships are identified, 
while the node distance limits in the construction phase control redundancy prevention. The 
asymmetric limits for opposite relationships (100m forward vs. 10m backward) reflect that vehicles 
traveling in opposite directions are more relevant when they are ahead rather than behind.

All parameters are configurable and can be adjusted based on the specific application 
requirements, road network characteristics, or desired graph density. The values shown in 
Table~\ref{tab:construction_parameters} represent default values that work well for urban traffic 
scenarios, but may need adjustment for highway scenarios (longer distances) or dense urban 
intersections (shorter distances).

%TODO: add examples

\subsection{Time based graph representations}
The graph creation strategy described above is assuming to create a graph for a single timestep. 
One possible extension is to create a graph which contains information for multiple timesteps. In 
order to create a time indexed graph, first the actor graphs for all timesteps are created. The 
joint set of all nodes across all timesteps is then determined. Node attributes of single timestep 
graphs are combined into a tensor of node attributes, introducing a new dimension for the time 
index. While the nodes are somehow time independent by this strategy, the edges are modeled as 
time dependent. Here it is utilized, that there can be multiple edges between the same pair of 
nodes, keeping potential edge attributes and adding an additional edge attribute to indicate the 
time stamp of the edge.

This extension allows the graph to easily represent changing states, e.g. the relationship between 
two vehicles changes from neighbor to leading vehicle, i.e. a cut in happened. One important 
special case for this type of processing is to handle actors, which are not present in all time 
steps, i.e. they appear and disappear during the time window. Especially if node attributes are 
used in the processing, a suitable imputation strategy is necessary to avoid non-equal dimensions 
of tensors for node attributes.

The methods derived in the following chapters can be applied to this extended graph representation 
as well, even though the focus in this paper will be on single timestep graphs.