### 2025-11-23 article/sections/existing_approaches.tex

\section{Existing coverage and analysis approaches}
\label{chapter:existing_approaches}

add Master thesis Johannes

There exist a large amount of literature on coverage analysis in the context of autonomous driving.
In the following, some of the most relevant approaches are discussed.

The PEGASUS project (\cite{pegasus2019method}) introduces a systematic, scenario-based methodology for the verification and 
validation of highly automated driving functions, addressing the impracticality of traditional 
distance-based testing approaches. The core of the method is a six-layer model used to 
systematically describe and structure the driving environment, encompassing factors from road geometry to weather conditions. 
This framework is particularly relevant for coverage analysis, as it facilitates a structured decomposition of the vast, 
continuous test space into discrete, manageable "logical scenarios". By systematically 
parameterizing and exploring these logical scenarios across simulation, proving ground, and field tests, 
the methodology aims to ensure the completeness of relevant test runs and provides a structured 
foundation for arguing that the automated system has been adequately tested across its entire operational design 
domain (ODD). This shift from random sampling to a structured, scenario-based approach allows 
for a more efficient and comprehensive assessment to generate evidence for a final safety argumentation

In a similar direction, \cite{deGelder2022ontology} focus on defining ontoligies for automated vehicles
in an object oriented manner, with the intention to have a clearly defined and implementable
structures for scenarios, delivering a clear linkage to coverage arguments.

The authors of (\cite{Ries2021traj_clustering}) address the challenge of validating automated driving systems 
in complex urban environments by proposing a trajectory-based clustering method for real-world driving data. 
They extract motion trajectories and associated features from urban driving recordings, apply unsupervised 
clustering to group similar driving behaviours/scenes, then analyze the resulting clusters to reveal common 
scenario types and redundancies. Their approach enables structuring the enormous scenario space, 
supporting more efficient test-set generation and scenario selection for automated vehicle verification. While 
not explicitly focused on coverage analysis, their approach is a good starting point for coverage analysis.

In \cite{Ulbrich2015scene}, the authors identify that key concepts in automated driving—namely scene, 
situation, and scenario—are inconsistently defined in the literature, which complicates the 
development, testing and validation of driving-automation modules. They propose clear 
definitions: a scene is a snapshot of the environment including dynamic and static 
elements plus actors' self-representations; a situation is the set of all circumstances relevant for 
behaviour decision at a given moment, derived from the scene but reflecting the actor's goals and 
values; and a scenario is a temporal development of several scenes in sequence, involving 
actions/events and goals/values. The paper also provides example implementations of these 
definitions in the context of automated-vehicle systems and how they interface with perception, 
planning and control, and testing.

Another important reference is the SAE International recommended practice \cite{ORAD2021taxonomy}, which 
provides many foundational terms in the context of autonomous driving.
It provides a functional taxonomy and clear definitions for key terms related to 
driving automation systems (DAS) used in on-road motor vehicles. The document defines six 
levels of driving automation (Levels 0 through 5) based on the role of three primary actors — 
the human driver, the driving automation system (DAS), and other vehicle systems/components. It introduces 
important related terms such as the Dynamic Driving Task (DDT), Operational Design 
Domain (ODD), Automated Driving System (ADS), and Vehicle Motion Control, 
among others. The 2021-04 revision (superseding the 2018 version) was developed in 
collaboration with ISO/TC 204/WG14 to harmonise global terminology and improve clarity for multi-discipline audiences 
(engineering, legal, media). The document emphasises that it is descriptive (not normative); 
it does not prescribe specifications or impose performance requirements for DAS.

The paper \cite{DBLP:journals/corr/abs-1801-08598} proposes a scenario-based framework for developing and 
validating automated-driving systems across different development phases. It 
introduces three abstraction levels of scenarios—functional, logical, and 
concrete—and discusses how these can be transformed for use in testing. The authors note that 
existing parameter-selection methods, such as equivalence-class or combinatorial testing, lack a 
systematic way to determine meaningful test coverage, highlighting coverage analysis as 
an open challenge in scenario-based validation.

In \cite{Ammann_Offutt_2008}, a standard textbook on software testing although not specifically focused on 
autonomous driving, the authors provide an introduction to software testing, which is a 
good starting point for coverage analysis. They introduce the concept of coverage and discuss the 
different types of coverage, such as statement coverage, branch coverage, condition coverage, 
and decision coverage. They also discuss the different techniques for measuring coverage, such as 
branch coverage, condition coverage, and decision coverage.

A much more focussed example of coverage analysis in the context of autonomous driving is provided 
by \cite{foretellix2019} in a blog 
post, discussing topics like items, parameters and 
coverage buckets and performance metrics like time to collision and error collections.

The authors of (\cite{wachenfeld2016release}) investigate the significant challenge of safety validation and 
production release for fully autonomous vehicles, positing that established testing concepts are insufficient 
as they fundamentally rely on the human driver's ability to intervene as a safety backup. 
The authors introduce the "approval-trap," a statistical argument demonstrating the unfeasibility of proving 
superior safety through real-world driving, which would necessitate billions of test kilometers. 
This validation gap is presented as a fundamental problem of coverage analysis, where the
technical system must now demonstrably cover the vast operational domain previously managed by the 
human. The paper concludes that overcoming this challenge requires a paradigm shift toward new 
test case generation methodologies, such as critical scenario identification, and the extensive
 use of validated simulation-based tools to achieve sufficient test coverage efficiently.

Well known standards for coverage analysis in the context of autonomous driving are the 
ISO 21448 (\cite{iso21448}) and the UL 4600 (\cite{ul4600}).
The ISO 21448 is a standard for the safety of the intended functionality of road vehicles, 
including automated driving systems.
It defines a framework for coverage analysis, including the definition of coverage criteria and 
the measurement of coverage.
The UL 4600 is a standard for the safety of autonomous products, including automated driving systems.
It defines a framework for coverage analysis, including the definition of coverage criteria and the 
measurement of coverage.