### 2025-11-23 article/sections/summary.tex

\section{Summary}

\begin{itemize}
    \item A method for using traffic scenes represented as graphs for coverage analysis has been developed in this paper.
    \item Starting from defining rules for the actor graph construction, the resulting graphs are then used by 2 different approaches for coverage analysis.
    \item Firstly, a non-AI approach is used where a set of archetypes is defined (which is done here manually but in principle could be done automatically as well).
    \item Then it is checked, which archetypes are present in a traffic scene and how often.
    \item Secondly, a graph embedding approach is used to create a vector space of the traffic scenes and to check if the target distribution is met by the test distribution in the embedded space.
    \item The results for this approach are shown in Section \ref{chapter:application}, being applied to Carla and Argoverse 2.0 data.
    \item The main advantage of this approach over traditional coverage analysis approaches is that it very easily scales to a large number of traffic scenarios archetypes and different driving situations.
    \item Besides the definition of the actor graph, no specific handling of different types of traffic situations is necessary.
    \item In other approaches (cite TNO Streetwise) it is necessary to manually define for each scenario type how it is defined. Here the only thing necessary is to define the actor graph construction rules as well as potentially the archetypes as graphs.
    \item Also, the presented approach easily allows for different number of actors in the traffic scene without excluding actors or only focussing on a subset of actors.
    \item Next steps in the same line of this research will be to further optimize the actor graph to include several time steps into one graph. 
    \item Also, automatically extracting archetypes from real world traffic observation data will be investigated.
    \item The source code for this research is available at \url{https://github.com/tmuehlen80/graph_coverage}.
\end{itemize}