### 2025-12-04 article/article.tex

\documentclass{llncs}
% \usepackage{cite}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{svg}
\usepackage{subcaption}
\usepackage{amsmath}

\title{A Graph-based Framework for Coverage Analysis in Autonomous Driving}

\author{Thomas Mühlenstädt and Marius Bause}
\date{\today}

\begin{document}

\maketitle

\input{sections/abstract}
\input{sections/introduction}
\input{sections/existing_approaches}
\input{sections/defining_traffic_scene_graph}
\input{sections/analysing_traffic_scene}
\input{sections/create_subgraphs}
\input{sections/graph_embeddings}
\input{sections/application}
\input{sections/summary}

\bibliographystyle{splncs04}
\bibliography{lit}

\end{document}